On The Impact of Atoms of Confusion in JavaScript Code

### 2021-09-27 5_Threats.tex

\textbf{Conclusion validity.} In the context of our survey study, we apply different statistical tests appropriate for the cases where data was categorical (correctness, Chi-square test of independence) and continuous (time, Mann-Whitney U test). Furthermore, besides reporting p-values, we also report effect size measures appropriate for each scenario (odds-ratio and Cliff's delta) and apply a p-value correction technique to avoid the multiple testing problem. Finally, it could be argued that the size of the samples is insufficient to make conclusions for some of the atoms, a common problem in Empirical Software Engineering. We estimate the sample size for each one of the atom candidates, considering that each one had a different number of samples (Table~\ref{tab:difference-correctness}). To that end, we use the $\phi$ measure of effect size for each atom (based on the Chi-square statistic), the standard $\alpha$ coefficient of 0.05, and set the expected statistical power to 0.8, as usually employed in the literature~\cite{Ellis:2010:EGE}. As a result of this analysis, we discover that for all but one of the atoms where we had a statistically significant result for correctness, Implicit Predicate, the sample size is sufficient.
% For the ones where we did not find statistically significant results for correctness, in fact the sample sizes
% we have analyzed are insufficient for such small effect sizes. This indicates that further studies on these
% atoms are required, due to the likelihood of type II errors.

%One threat to validity we have not properly addressed is lack of power, a common problem in Software Engineering research.

@@

\textbf{Construct validity.} We use correctness as a proxy for program comprehension. Furthermore, we use predicted outcomes of small programs as a measure of correctness. As discussed elsewhere~\cite{Oliveira:2020:ECR}, this approach is a test of the developers' ability to trace programs. Although this is a common approach in studies about atoms of confusion~\cite{TheEyesDoNotLie,Langhout:2021:ACJ,DBLP:conf/sigsoft/GopsteinIYDZYC17}, other measures of correctness could have been employed, potentially yielding different results. As a complement to correctness in the survey, we have measured the time it took for the participants to correctly predict the outcome of the code snippets (either with or without atom candidates). Furthermore, we have  asked the interviewees about their preferences when comparing confusing and clean versions of the programs.

@@

\textbf{External validity.} Although our results suggest that the selected idioms and code constructs may lead to confusion for small code snippets, it is not clear whether that result extrapolates to other scenarios. Even though it is likely that in larger code bases the confusion induced by these constructs may be even greater, the existence of additional context may mitigate this effect. Another possible threat to the generalizability of our results, in particular for the survey and interviews, lies in the fact that the analyzed atoms may rarely occur in real JavaScript code bases. To mitigate that threat, we have analyzed 72 popular open source JavaScript repositories and found out that most of them are common, occurring at least once per 1,000 lines of code. Only Arithmetic as Logic and Comma Operator occur less frequently than once per 10,000 lines of code. 
% \castor{A forma como os átomos foram detectados é uma ameaça que não discutimos.}

>>>>

\textbf{Conclusion validity.} 
%Threats to conclusion validity often come from inappropriate use of statistics. 
In the context of our survey study, we apply different non-parametric statistical tests appropriate for the cases where data was categorical (correctness, Chi-square test of independence) and continuous (time, Mann-Whitney U test). Furthermore, besides reporting p-values, we also report effect size measures appropriate for each scenario (odds-ratio and Cliff's delta) and apply a p-value correction technique to avoid the multiple testing problem. Finally, it could be argued that the size of the samples is insufficient to make conclusions for some of the atoms, a common problem in Empirical Software Engineering. We estimate the sample size for each one of the atom candidates, considering that each one had a different number of samples (Table~\ref{tab:difference-correctness}). To that end, we use the $\phi$ measure of effect size for each atom (based on the Chi-square statistic), the standard $\alpha$ coefficient of 0.05, and set the expected statistical power to 0.8, as usually employed in the literature~\cite{Ellis:2010:EGE}. We find out that the sample size is sufficient for all but one of the atoms where we had a statistically significant result for correctness, Implicit Predicate. 
% For the ones where we did not find statistically significant results for correctness, in fact the sample sizes
% we have analyzed are insufficient for such small effect sizes. This indicates that further studies on these
% atoms are required, due to the likelihood of type II errors.

@@

\textbf{Construct validity.} We use correctness as a proxy for program comprehension and predicted outcomes of small programs as a measure of correctness. As discussed elsewhere~\cite{Oliveira:2020:ECR}, this approach is a test of the developers' ability to trace programs. Although this is a common approach in studies about atoms of confusion~\cite{TheEyesDoNotLie,Langhout:2021:ACJ,DBLP:conf/sigsoft/GopsteinIYDZYC17}, other measures of correctness could have been employed, potentially yielding different results. As a complement to correctness in the survey, we have measured the time it took for the participants to correctly predict the outcome of the code snippets (either with or without atom candidates). Furthermore, we have  asked the interviewees about their preferences when comparing confusing and clean versions of the programs.

@@

\textbf{External validity.} Our results suggest that the selected idioms and code constructs may lead to confusion for small code snippets, but it is not clear if that result extrapolates to other scenarios. Even though it is likely that in larger code bases the confusion induced by these constructs may be even greater, the existence of additional context may mitigate this effect. Another possible threat to the generalizability of our results, in particular for the survey and interviews, lies in the fact that the analyzed atoms may rarely occur in real JavaScript code bases. To mitigate that threat, we have analyzed 72 popular open source JavaScript repositories and found out that most of them are common, occurring at least once per 1,000 lines of code. Only Arithmetic as Logic and Comma Operator occur less frequently than once per 10,000 lines of code. 
% \castor{A forma como os átomos foram detectados é uma ameaça que não discutimos.}
